% ..\ideas\minority_attack
% ..\ideas\breakthrough_f3_e4

The Benoni structure comes typically from Benoni defence. It can also come from English opening (reversed), Ruy Lopez or Italian after Black gives up the center with exd4, c5 and White answers with d5. 

\newchessgame[
id=A,
moveid=1w,
setwhite={pa2, pb2, pd5, pe4, pf2, pg2, ph2},
addblack={pa7, pb7, pc5, pd6, pf7, pg6, ph7}]
\chessboard

The exchange of White's c-pawn for Black's e-pawn leaves White with a pawn majority in the centre and Black with one on the queenside. This asymmetry suggests that White will try to play on the kingside and in the centre, while Black will seek counterplay on the queenside. However, this simplistic generalization does not hold in many cases—depending on how the pieces are arranged, either side may be able to fight back on the flank where they are theoretically weaker.


\minisec{White Plans \footnote{\url{https://en.wikipedia.org/wiki/Modern_Benoni}}}
\begin{itemize}
    \item{King side play}

    The central pawn majority is White's main positional trump in the Modern Benoni. By staking out an advantage in space on the kingside, it allows White to develop an initiative on that side of the board. The most important tool in White's arsenal is the e4-e5 pawn advance, which can open up lines and squares for the white pieces, and result in the creation of a passed d-pawn if Black answers with ...dxe5.
    \item{Queen side play}

    When Black prepares the ...b7-b5 pawn break with ...a6, White usually tries to hinder it by playing a2-a4, even though this advance weakens the b4-square. As a further deterrent to Black's queenside expansion, White often moves the knight on f3 to c4 via d2. With the knight on c4, Black's ...b7-b5 break may be met by axb5 followed by \symknight{}a5, when the arrival of a white knight on c6 could severely disrupt Black's position. The knight on c4 also attacks Black's backward pawn on d6, and White can often increase the pressure on this pawn by playing Bf4 or Nb5. The strength of White's knight on c4 often induces Black to exchange it off: typical ways of doing so are ...\symknight{}b6, ...\symknight{}e5, or ...b7-b6 followed by ...\symbishop{}a6.
\end{itemize}    

\minisec{Black's Key Ideas}
\begin{itemize}
    \item{Queen side play}

    Queen side pawn rush supported by the g7 bishop to create a pass pawn. 

    \item{c5-c4}

    Push the pawn to make room for the d7 knight to create some tactical chance.
    \item{King side play}

    The half-open e-file gives Black a certain degree of influence over the kingside. A rook on e8 puts pressure on White's e-pawn and restrains it from advancing. Tactics involving ...\symknight{} xe4 are not uncommon.
\end{itemize} 



